Figure Extra

### 2022-06-27 Figures/Figures_nonTCGA.Rmd
Fix corrplot for GSE72056

---
title: "Figure 2 and figure 2 supplementary material"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

@@

################################################################################
benchmark_corrplots <- function(deconv_data, true_estimates){
  # build the data
  true_subset_ <- true_estimates %>%  na.exclude() %>%
    arrange(factor(ParticipantBarcode, levels = ParticipantBarcode))

  Comm_samples <- Reduce(intersect, list(deconv_data$Samples,
                                         true_subset_$ParticipantBarcode))

  true_subset <-dplyr::filter(true_subset_, ParticipantBarcode %in% Comm_samples) %>% 
    arrange(factor(ParticipantBarcode, levels = Comm_samples))

  deconv_subset <- dplyr::filter(deconv_data, Samples %in% Comm_samples) %>% 
    arrange(factor(Samples, levels = Comm_samples))

  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict)
  identical(true_subset$ParticipantBarcode, deconv_subset_summary$Samples) %>% stopifnot()

  to_corr <- bind_cols(true_subset, deconv_subset_summary) %>% dplyr::select(-c(Samples))
  ct_corrplots <- list()

    for (st in unique(celltypes)){
      deconv_ct_names <- dplyr::filter(to_corr, Study == st) %>%
        # segregate by ct
        dplyr::select(matches(ct)) %>%
        formatted_cors(cor.stat = "pearson") %>%
        # reduce to corrs just to ground truth
        dplyr::filter(measure1 == all_of(ct), measure2 != all_of(ct)) %>%
        # rename ground truth as cancer type
        dplyr::mutate(measure1 = st)
      #add to the rest of cancer type correlations
      store_corrs <- add_row(store_corrs , deconv_ct_names)
    }

    # Plot
    store_corrs <- dplyr::mutate(store_corrs, measure2=stringr::str_remove_all(measure2, paste(substring_remove, collapse = "|")))

    order <- store_corrs %>% 
      drop_na(r) %>% # Epidish_CBSX_LM22 dont work for UVM in Neutrophils and mean dont work with NANs
      group_by(measure2) %>%
      dplyr::summarize(r = mean(r)) %>% 
      arrange(desc(r)) %>% 
      mutate(measure1 = "Average",
             p = NA,
             sig_p = TRUE,
             p_if_sig = NA,
             r_if_sig = r ) %>%
      dplyr::relocate(measure1, .before = measure2)

@@

    ct_corrplots[[ct]] <- add_row(store_corrs, order) %>% dplyr::mutate(measure2 = factor(measure2, levels = order$measure2)) %>% dplyr::arrange(factor(measure2, levels = order$measure2)) %>%
    ggplot(aes(measure2, measure1, fill=r, label=round(r_if_sig,2))) +
    geom_tile() +
      labs(x = NULL, y = NULL, fill = "Spearman's\nCorrelation", title= ct, 
         subtitle="Only significant correlation coefficients shown (95% I.C.)") +
    scale_fill_gradient2(mid="#FBFEF9",low="#0C6291",high="#A63446", limits=c(-1,1)) +
    geom_text() +
    theme_classic() +
    scale_x_discrete(expand=c(0,0)) +
    scale_y_discrete(expand=c(0,0), labels= function(x) highlight(x, "Average", "red")) +
    rotate_x_text(angle = 45) + theme(axis.text.y=element_markdown())

@@

# Data loading

```{r}
deconv <- read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/Projects/CRCT/21 - Epigenomics and network modelling of heterogeneity in immuno-oncology/Ting Xie/2021/CIBERSORTx/final_example_signatures/all_deconvolutions/all_deconvolutions_Melanoma_GSE72056_no_malighant.txt")

@@

# Figure
## B) 3 Celltypes ridge plots
### Create aggregation dictionaries to compare with Lymphocites, Macrophages and Neutrophils FACs

```{r}
# Lists of celltypes of each signature
## signature dictionary construction (can use . as regexp!)
## uncoment print of functions to check that you detect what you want.
## non matching signatures/methods (all 0) are removed.
signature_dict <- list()
signature_dict[["BPRNACan"]] <- list(B = c("B"),
                                     CAFs = c(""),
                                     CD4 = c("CD4"),
                                     CD8 = c("CD8"),
                                     Endothelial = c(""),
                                     Macrophages = c("M0", "M1", "M2"),
                                     NK = c("NK"),
                                     Th = c(""),
                                     Treg = c(""),
                                     Cancer = c("cancer"),
                                     Other = c("Mono", "Neu"))

@@

signature_dict[["CBSX_Fig2ab-NSCLC_PBMCs_scRNAseq_sigmatrix"]] <- list(B = c("B.cells"),
                                     CAFs = c(""),
                                     CD4 = c("T.cells.CD4"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c(""),
                                     Macrophages = c(""),
                                     NK = c("NK.cells"),
                                     Th = c(""),
                                     Treg = c(""),
                                     Cancer = c(""),
                                     Other = c("Monocytes","NKT.cells"))

signature_dict[["CBSX_LM22"]] <- list(B = c("B.cells.naive", "B.cells.memory", "Plasma.cells"),
                                     CAFs = c(""),
                                     CD4 = c("T.cells.CD4.naive", "T.cells.CD4.memory.resting", "T.cells.CD4.memory.activated"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c(""),
                                     Macrophages = c("Macrophages.M0", "Macrophages.M1", "Macrophages.M2"),
                                     NK = c("NK.cells.resting", "NK.cells.activated"),
                                     Th = c("T.cells.follicular.helper", "T.cells.gamma.delta"),
                                     Treg = c("T.cells.regulatory..Tregs."),
                                     Cancer = c(""),
                                     Other = c("Monocytes", "Dendritic.cells.resting", "Dendritic.cells.activated", "Mast.cells.resting", "Mast.cells.activated", "Eosinophils", "Neutrophils"))

signature_dict[["CBSX_scRNA-Seq_melanoma_Tirosh_sigmatrix_SuppFig_3-b"]] <- list(B = c("B.cells"),
                                     CAFs = c("CAF"),
                                     CD4 = c("T.cells.CD4"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c("Endothelial.cells"),
                                     Macrophages = c("Macrophages"),
                                     NK = c("NK.cells"),
                                     Th = c(""),
                                     Treg = c(""),
                                     Cancer = c("Malignant"),
                                     Other = c(""))

signature_dict[["CBSX_sigmatrix_HNSCC_Fig2cd"]] <- list(B = c("B.cells"),
                                     CAFs = c("CAFs"),
                                     CD4 = c("T.cells.CD4"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c("Endothelial.cells"),
                                     Macrophages = c("Macrophages"),
                                     NK = c(""),
                                     Th = c(""),
                                     Treg = c(""),
                                     Cancer = c("Malignant.cells"),
                                     Other = c("Dendritic.cells", "Mast.cells", "Myocytes"))

signature_dict[["Quantiseq"]] <- list(B = c("B.cell"),
                                     CAFs = c(""),
                                     CD4 = c("T.cell.CD4+.(non-regulatory)"),
                                     CD8 = c("T.cell.CD8+"),
                                     Endothelial = c(""),
                                     Macrophages = c("Macrophage.M1", "Macrophage.M2"),
                                     NK = c("NK.cell"),
                                     Th = c(""),
                                     Treg = c("T.cell.regulatory(Tregs)"),
                                     Cancer = c(""),
                                     Other = c("Monocyte", "Neutrophil", "Myeloid.dendritic.cell", "uncharacterized.cell"))

# signature_dict[["CCLE_TIL10"]] <- list(B = c("B"),
#                                      CAFs = c(""),
#                                      CD4 = c(""),
#                                      CD8 = c(""),
#                                      Endothelial = c(""),
#                                      Macrophages = c(""),
#                                      NK = c(""),
#                                      Th = c(""),
#                                      Treg = c(""),
#                                      Cancer = c(""),
#                                      Other = c(""))

# signature_dict[[""]] <- list(B = c(""),
#                                      CAFs = c(""),
#                                      CD4 = c(""),
#                                      CD8 = c(""),
#                                      Endothelial = c(""),
#                                      Macrophages = c(""),
#                                      NK = c(""),
#                                      Th = c(""),
#                                      Treg = c(""),
#                                      Cancer = c(""),
#                                      Other = c(""))

```
## Summary Corrplot

```{r}
corrplots <- benchmark_corrplots(deconv_data = deconv, true_estimates = ground_truth)
corrplots$Macrophages <- corrplots$Macrophages + rremove("y.text")
corrplots$Neutrophils <- corrplots$Neutrophils + rremove("y.text")
summary_corrplot  <- ggarrange(plotlist = corrplots, ncol = 3, common.legend = T, widths = c(0.35, 0.35, 0.25) , legend = "bottom")

ggsave("Fig2_summary_corrplot.pdf", dpi = "print", width = 250, height = 100, units = "mm", scale = 2.5)
```

>>>>

---
title: "Figure Extra"
output:
  pdf_document: default
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

@@

################################################################################
# This function returns the Lymphocytes, Macrophages and Neutrophils summarized
# of a given method/signature.
# Works inside of ridgeplot_parse. it takes:
# -  'deconv_data': contains the deconvolution data to index and summarize. 
#
# -  'signature dictionary': contains the exact names of the method
#                           (if it works with one signature) or the signature
#                           (if it works with multiple methods). 
#
# - iterator: string to subscript the signatures_dict with either signature
#             or methods names depending where is it called
#
# - prefix: string to leave just celltype names when substracted to the columns
#             of deconv_data of the specified method (or method + signature).
#
# - celltype: name of the celltype to sumarize "Lymphocites", "Macrophages" or "Neutrophils"

ridgeplot_parse <- function(deconv_data, signatures_dictionary, key){

    results <- tibble(Samples = deconv_data[[key]])

    names_table <- deconv_data %>%
      dplyr::select(-any_of(key)) %>%
      names() %>%
      str_split_fixed("_", 2) %>%
      as_tibble(.name_repair = "unique")

    for (method in unique(names_table[,1,drop=T])){

      if (method %in% c("Epidish",
                        "DeconRNASeq",
                        "CBSX")){

        for (signature in names(signatures_dictionary)){ # iterates through all the signatures (the ones not present for the method will yield all 0)
          to_match <- paste(method, signature, sep = "_")

          for (celltype in names(signatures_dictionary[[signature]])){
            if (length(signatures_dictionary[[signature]][[celltype]]) > 0){
              name_of_summary <- paste(to_match, celltype, sep = "_")
              columns_to_summarize <- paste(to_match, signatures_dictionary[[signature]][[celltype]], sep = ".")

              print(paste0("Truth: ", name_of_summary))
              deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% colnames() %>% paste0("Est: ",.) %>% print()
              cat("-------------------------\n")
              deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% rowSums(.) -> results[name_of_summary]
            }
          }
    }
      } else{

      to_match <- method

      for (celltype in names(signatures_dictionary[[method]])){
        if (length(signatures_dictionary[[method]][[celltype]]) > 0){
            name_of_summary <- paste(to_match, celltype, sep = "_")
            columns_to_summarize <- paste(to_match, signatures_dictionary[[method]][[celltype]], sep = ".")

            deconv_data %>% dplyr::select(matches(columns_to_summarize)) %>% rowSums(.) -> results[name_of_summary]
        }
      }
    }
    }
    results %>% dplyr::select(-Samples) %>% dplyr::select(which(!colSums(., na.rm=TRUE) %in% 0)) %>% bind_cols(Samples = results$Samples,.)
}
################################################################################

################################################################################
benchmark_corrplots <- function(deconv_data, true_estimates, key, title){
  # build the data
  true_subset_ <- true_estimates %>%  na.exclude() %>%
    arrange(factor(get(key), levels = get(key)))

  common <- Reduce(intersect, list(deconv_data[[key]],
                   true_subset_[[key]]))

  true_subset <- dplyr::filter(true_subset_, get(key) %in% common) %>% 
    arrange(factor(get(key), levels = common))

  deconv_subset <- dplyr::filter(deconv_data, get(key) %in% common) %>% 
    arrange(factor(get(key), levels = common))

  deconv_subset_summary <- ridgeplot_parse(deconv_data = deconv_subset, signatures_dictionary = signature_dict, key = key)
  identical(true_subset[[key]], deconv_subset_summary[[key]]) %>% stopifnot()

  to_corr <- inner_join(true_subset, deconv_subset_summary) %>% dplyr::select(-any_of(key))
  ct_corrplots <- list()
  store_corrs <- tibble(measure1 = character(),
                          measure2 = character(),
                          r = numeric(),
                          n = numeric(),
                          p = numeric(),
                          sig_p = logical(),
                          p_if_sig = numeric(),
                          r_if_sig = numeric())

    for (ct in unique(dplyr::select(true_estimates, -any_of(key)) %>% names())){

      substring_remove <- c(paste0("_",ct,"$"),
      "_Tirosh_sigmatrix_SuppFig_3.b",
      "_scRNA.Seq",
      "Fig2ab.",
      "_PBMCs_scRNAseq_sigmatrix",
      "sigmatrix_",
      "_Fig2cd")

      deconv_ct_names <- to_corr %>%
        # segregate by ct
        dplyr::select(matches(paste0(ct, "$"))) %>%
        formatted_cors(cor.stat = "pearson") %>%
        # reduce to corrs just to ground truth
        dplyr::filter(measure1 == all_of(ct), measure2 != all_of(ct)) %>%
        # rename ground truth as cancer type
        dplyr::mutate(measure2 = str_remove_all(measure2,  paste(substring_remove, collapse = "|")))
      #add to the rest of cancer type correlations
      store_corrs <- add_row(store_corrs , deconv_ct_names)
    }

    # Plot

    order <- store_corrs %>% 
      drop_na(r) %>% 
      group_by(measure2) %>%
      dplyr::summarize(r = mean(r)) %>% 
      arrange(desc(r)) %>% 
      mutate(measure1 = "Average",
             p = NA,
             sig_p = TRUE,
             p_if_sig = NA,
             r_if_sig = r ) %>%
      dplyr::relocate(measure1, .before = measure2)

@@

    ct_corrplots[[ct]] <- add_row(store_corrs, order) %>% dplyr::mutate(measure2 = factor(measure2, levels = order$measure2)) %>% dplyr::arrange(factor(measure2, levels = order$measure2)) %>%
    ggplot(aes(measure1, measure2, fill=r, label=round(r_if_sig,2))) +
    geom_tile() +
      labs(x = NULL, y = NULL, fill = "Spearman's\nCorrelation", title= title, 
         subtitle="Only significant correlation coefficients shown (95% I.C.)") +
    scale_fill_gradient2(mid="#FBFEF9",low="#0C6291",high="#A63446", limits=c(-1,1)) +
    geom_text() +
    theme_classic() +
    scale_x_discrete(expand=c(0,0), labels= function(x) highlight(x, "Average", "red")) +
    scale_y_discrete(expand=c(0,0), labels= function(x) highlight(x, "Average", "red")) +
    rotate_x_text(angle = 45) + theme(axis.text.x=element_markdown()) + theme(axis.text.y=element_markdown())

@@

# GSE72056
## Data loading
```{r}
deconv <- read_tsv("/mnt/SERVER-CRCT-STORAGE/CRCT21/Private/Miguel/Projects/CRCT/21 - Epigenomics and network modelling of heterogeneity in immuno-oncology/Ting Xie/2021/CIBERSORTx/final_example_signatures/all_deconvolutions/all_deconvolutions_Melanoma_GSE72056_no_malighant.txt") %>% dplyr::rename_with(~str_replace(.,"__", "_CBSX_"))  %>% dplyr::rename_with(~str_replace(.,"__", "_")) #! fix naming of deconv results

@@

## Signature Dictionary construction

```{r}
# Lists of celltypes of each signature
## signature dictionary construction (can use . as regexp!)
## uncoment print of functions to check that you detect what you want.
## non matching signatures/methods (all 0) are removed.
signature_dict <- list()
signature_dict[["BPRNACan"]] <- list(B = c("B"),
                                     CAFs = c(),
                                     CD4 = c("CD4"),
                                     CD8 = c("CD8"),
                                     Endothelial = c(),
                                     Macrophages = c("M0", "M1", "M2"),
                                     NK = c("NK"),
                                     Th = c(),
                                     Treg = c(),
                                     Cancer = c("cancer"),
                                     Other = c("Mono", "Neu"))

@@

signature_dict[["CBSX_Fig2ab-NSCLC_PBMCs_scRNAseq_sigmatrix"]] <- list(B = c("B.cells"),
                                     CAFs = c(),
                                     CD4 = c("T.cells.CD4"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c(),
                                     Macrophages = c(),
                                     NK = c("NK.cells"),
                                     Th = c(),
                                     Treg = c(),
                                     Cancer = c(),
                                     Other = c("Monocytes","NKT.cells"))

signature_dict[["CBSX_LM22"]] <- list(B = c("B.cells.naive", "B.cells.memory", "Plasma.cells"),
                                     CAFs = c(),
                                     CD4 = c("T.cells.CD4.naive", "T.cells.CD4.memory.resting", "T.cells.CD4.memory.activated"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c(),
                                     Macrophages = c("Macrophages.M0", "Macrophages.M1", "Macrophages.M2"),
                                     NK = c("NK.cells.resting", "NK.cells.activated"),
                                     Th = c("T.cells.follicular.helper", "T.cells.gamma.delta"),
                                     Treg = c("T.cells.regulatory..Tregs."),
                                     Cancer = c(),
                                     Other = c("Monocytes", "Dendritic.cells.resting", "Dendritic.cells.activated", "Mast.cells.resting", "Mast.cells.activated", "Eosinophils", "Neutrophils"))

signature_dict[["CBSX_scRNA-Seq_melanoma_Tirosh_sigmatrix_SuppFig_3-b"]] <- list(B = c("B.cells"),
                                     CAFs = c("CAF"),
                                     CD4 = c("T.cells.CD4"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c("Endothelial.cells"),
                                     Macrophages = c("Macrophages"),
                                     NK = c("NK.cells"),
                                     Th = c(),
                                     Treg = c(),
                                     Cancer = c("Malignant"),
                                     Other = c())

signature_dict[["CBSX_sigmatrix_HNSCC_Fig2cd"]] <- list(B = c("B.cells"),
                                     CAFs = c("CAFs"),
                                     CD4 = c("T.cells.CD4"),
                                     CD8 = c("T.cells.CD8"),
                                     Endothelial = c("Endothelial.cells"),
                                     Macrophages = c("Macrophages"),
                                     NK = c(),
                                     Th = c(),
                                     Treg = c(),
                                     Cancer = c("Malignant.cells"),
                                     Other = c("Dendritic.cells", "Mast.cells", "Myocytes"))

signature_dict[["Quantiseq"]] <- list(B = c("B.cell"),
                                     CAFs = c(),
                                     CD4 = c("T.cell.CD4+.(non-regulatory)"),
                                     CD8 = c("T.cell.CD8+"),
                                     Endothelial = c(),
                                     Macrophages = c("Macrophage.M1", "Macrophage.M2"),
                                     NK = c("NK.cell"),
                                     Th = c(),
                                     Treg = c("T.cell.regulatory(Tregs)"),
                                     Cancer = c(),
                                     Other = c("Monocyte", "Neutrophil", "Myeloid.dendritic.cell", "uncharacterized.cell"))

# signature_dict[["CCLE_TIL10"]] <- list(B = c(),
#                                      CAFs = c(),
#                                      CD4 = c(),
#                                      CD8 = c(),
#                                      Endothelial = c(),
#                                      Macrophages = c(),
#                                      NK = c(),
#                                      Th = c(),
#                                      Treg = c(),
#                                      Cancer = c(),
#                                      Other = c())

# signature_dict[[""]] <- list(B = c(),
#                                      CAFs = c(),
#                                      CD4 = c(),
#                                      CD8 = c(),
#                                      Endothelial = c(),
#                                      Macrophages = c(),
#                                      NK = c(),
#                                      Th = c(),
#                                      Treg = c(),
#                                      Cancer = c(),
#                                      Other = c())

```
## Corrplot

```{r}
benchmark_corrplots(deconv_data = deconv, true_estimates = ground_truth,  key = "Samples", title = "GSE72056")
# corrplots$Macrophages <- corrplots$Macrophages + rremove("y.text")
# corrplots$Neutrophils <- corrplots$Neutrophils + rremove("y.text")
# summary_corrplot  <- ggarrange(plotlist = corrplots, ncol = 3, common.legend = T, widths = c(0.35, 0.35, 0.25) , legend = "bottom")
# 
# ggsave("Fig2_summary_corrplot.pdf", dpi = "print", width = 250, height = 100, units = "mm", scale = 2.5)
```
